### 2018-02-20 submission/manuscript.Rmd
update methods

***Human and Mock Samples:***
A single fecal sample was obtained from 4 individuals who were part of the Enterics Research Investigational Network (ERIN). The processing and storage of these samples have been published previously [@erin_seekatz_2016]. Other than confirmation that none of these individuals had a *Clostridium difficle* infection clinical data and other types of meta data were not utilized or accessed for this study. All samples were extracted using the MOBIO^TM^ PowerMag Microbiome RNA/DNA extraction kit (now Qiagen, MD, USA). The ZymoBIOMICS^TM^ Microbial Community DNA Standard (Zymo, CA, USA) was used in this study and is made up of *Pseudomonas aeruginosa*, *Escherichia coli*, *Salmonella enterica*, *Lactobacillus fermentum*, *Enterococcus faecalis*, *Staphylococcus aureus*, *Listeria monocytogenes*, and *Bacillus subtilis* at equal genomic DNA abundance  (http://www.zymoresearch.com/microbiomics/microbial-standards/zymobiomics-microbial-community-standards).

***PCR Protocol:***
The five different high fidelity (HiFi) DNA polymerases that were tested included AccuPrime^TM^ (ThermoFisher, MA, USA), KAPA HIFI (Roche, IN, USA), Phusion (ThermoFisher, MA, USA), Platinum (ThermoFisher, MA, USA), and Q5 (New England Biolabs, MA, USA). The PCR cycle conditions for Platinum and Accuprime followed a previously published protocol [@protocol_Kozich_2013] (https://github.com/SchlossLab/MiSeq_WetLab_SOP/blob/master/MiSeq_WetLab_SOP_v4.md). The HiFi DNA polymerase activation time was 2 minutes, unless a different activation was specified. For Kappa and Q5, the protocol previously published by Gohl and colleagues was used [@taq_Gohl_2016]. For Phusion, the company defined conditions were used except for extension time, where the Accuprime and Platinum settings were used.

@@

***Sequence Processing:***
The mothur software program was utilized for all sequence processing steps [@mothur_schloss_2009]. Generally, the protocol followed what has been previously published [@protocol_Kozich_2013] (https://www.mothur.org/wiki/MiSeq_SOP). Two major differences from the stated protocol were the use of VSEARCH instead of UCHIME for chimera detection and the use of the OptiClust algorithm instead of average neighbor for Operational Taxonomic Unit (OTU) generation at 97% similarity [@vsearch_Rognes_2016; @uchime_Edgar_2011; @opticlust_Westcott_2017]. Sequence error was determined using the 'seq.error' command on mock samples after chimera removal and classification to the RDP to remove non-bacterial sequences [@mothur_schloss_2009; @vsearch_Rognes_2016; @rdp_Cole_2013].

@@

***Analysis Workflow:***
The total number of OTUs after sub-sampling was analyzed for both the fecal and mock community samples. Cycle dependent affects on Bray-Curtis indices were next assessed for the fecal samples looking at both overall cycle differences and within individual differences for the previous cycle (e.g. 20 versus 25, 25 versus 30, etc.). For these community based measures, the fecal samples were analyzed at 4 different sub-sampling sequence depths (1000, 5000, 10000, and 15000) while the mock community samples were analysed at 3 levels (1000, 5000, 10000). Next, based on these observations we wanted to analyze potential reasons for these differences. First, analysis of general sequence error rate, number of sequences with an error, and base substitution were assessed in the mock community for each DNA polymerase. After assessing these errors, the total number of chimeras was determined after sequence processing.

>>>>

***Human and Mock Samples:***
A single fecal sample was obtained from 4 individuals who were part of the Enterics Research Investigational Network (ERIN). The processing and storage of these samples have been previously published [@erin_seekatz_2016]. Other than confirmation that none of these individuals had a *Clostridium difficle* infection, clinical data and other types of meta data were not utilized or accessed for this study. All samples were extracted using the MOBIO^TM^ PowerMag Microbiome RNA/DNA extraction kit (now Qiagen, MD, USA). The ZymoBIOMICS^TM^ Microbial Community DNA Standard (Zymo, CA, USA) was used in this study and is made up of *Pseudomonas aeruginosa*, *Escherichia coli*, *Salmonella enterica*, *Lactobacillus fermentum*, *Enterococcus faecalis*, *Staphylococcus aureus*, *Listeria monocytogenes*, and *Bacillus subtilis* at equal genomic DNA abundance (http://www.zymoresearch.com/microbiomics/microbial-standards/zymobiomics-microbial-community-standards).

***PCR Protocol:***
The five different HiFi DNA polymerases that were tested included AccuPrime^TM^ (ThermoFisher, MA, USA), KAPA HIFI (Roche, IN, USA), Phusion (ThermoFisher, MA, USA), Platinum (ThermoFisher, MA, USA), and Q5 (New England Biolabs, MA, USA). The PCR cycle conditions for Platinum and Accuprime followed a previously published protocol [@protocol_Kozich_2013] (https://github.com/SchlossLab/MiSeq_WetLab_SOP/blob/master/MiSeq_WetLab_SOP_v4.md). The HiFi DNA polymerase activation time was 2 minutes, unless a different activation was specified. For Kappa and Q5, the protocol previously published by Gohl and colleagues was used [@taq_Gohl_2016]. For Phusion, the company defined conditions were used except for extension time, where the Accuprime and Platinum settings were used.

@@

***Sequence Processing:***
The mothur software program was utilized for all sequence processing steps [@mothur_schloss_2009]. Generally, the protocol followed what has been previously published [@protocol_Kozich_2013] (https://www.mothur.org/wiki/MiSeq_SOP). Two major differences from the stated protocol were the use of VSEARCH instead of UCHIME for chimera detection and the use of the OptiClust algorithm instead of average neighbor for OTU generation at 97% similarity [@vsearch_Rognes_2016; @uchime_Edgar_2011; @opticlust_Westcott_2017]. Sequence error was determined using the 'seq.error' command on mock samples after chimera removal and classification to the RDP to remove non-bacterial sequences [@mothur_schloss_2009; @vsearch_Rognes_2016; @rdp_Cole_2013].

***Analysis Workflow:***
The total number of OTUs after sub-sampling was analyzed for both the fecal and mock community samples. Cycle dependent affects on Bray-Curtis indices were next assessed for the fecal samples looking at both overall cycle differences and within individual differences for the previous cycle (e.g. 20 versus 25, 25 versus 30, etc.). For these community based measures, the fecal samples were analyzed at 4 different sub-sampling sequence depths (1000, 5000, 10000, and 15000) while the mock community samples were analysed at 3 levels (1000, 5000, 10000). Next, based on these observations we wanted to analyze potential reasons for these differences. First, analysis of general sequence error rate, number of sequences with an error, and base substitution were assessed in the mock community for each HiFi DNA polymerase. After assessing these errors, the total number of chimeras was determined after sequence processing and correlations to the number of OTUs also assessed.

@@

### 2018-02-20 submission/manuscript.tex
update methods

\textbf{\emph{Human and Mock Samples:}} A single fecal sample was
obtained from 4 individuals who were part of the Enterics Research
Investigational Network (ERIN). The processing and storage of these
samples have been published previously (Seekatz et al., 2016). Other
than confirmation that none of these individuals had a \emph{Clostridium
difficle} infection clinical data and other types of meta data were not
utilized or accessed for this study. All samples were extracted using
the MOBIO\textsuperscript{TM} PowerMag Microbiome RNA/DNA extraction kit
(now Qiagen, MD, USA). The ZymoBIOMICS\textsuperscript{TM} Microbial
Community DNA Standard (Zymo, CA, USA) was used in this study and is
made up of \emph{Pseudomonas aeruginosa}, \emph{Escherichia coli},
\emph{Salmonella enterica}, \emph{Lactobacillus fermentum},
\emph{Enterococcus faecalis}, \emph{Staphylococcus aureus},
\emph{Listeria monocytogenes}, and \emph{Bacillus subtilis} at equal
genomic DNA abundance
(\url{http://www.zymoresearch.com/microbiomics/microbial-standards/zymobiomics-microbial-community-standards}).

\textbf{\emph{PCR Protocol:}} The five different high fidelity (HiFi)
DNA polymerases that were tested included AccuPrime\textsuperscript{TM}
(ThermoFisher, MA, USA), KAPA HIFI (Roche, IN, USA), Phusion
(ThermoFisher, MA, USA), Platinum (ThermoFisher, MA, USA), and Q5 (New
England Biolabs, MA, USA). The PCR cycle conditions for Platinum and
Accuprime followed a previously published protocol (Kozich et al., 2013)
(\url{https://github.com/SchlossLab/MiSeq_WetLab_SOP/blob/master/MiSeq_WetLab_SOP_v4.md}).
The HiFi DNA polymerase activation time was 2 minutes, unless a
different activation was specified. For Kappa and Q5, the protocol
previously published by Gohl and colleagues was used (Gohl et al.,
2016). For Phusion, the company defined conditions were used except for
extension time, where the Accuprime and Platinum settings were used.

@@

\textbf{\emph{Sequence Processing:}} The mothur software program was
utilized for all sequence processing steps (Schloss et al., 2009).
Generally, the protocol followed what has been previously published
(Kozich et al., 2013) (\url{https://www.mothur.org/wiki/MiSeq_SOP}). Two
major differences from the stated protocol were the use of VSEARCH
instead of UCHIME for chimera detection and the use of the OptiClust
algorithm instead of average neighbor for Operational Taxonomic Unit
(OTU) generation at 97\% similarity (Edgar et al., 2011; Rognes et al.,
2016; Westcott \& Schloss, 2017). Sequence error was determined using
the `seq.error' command on mock samples after chimera removal and
classification to the RDP to remove non-bacterial sequences (Schloss et
al., 2009; Cole et al., 2013; Rognes et al., 2016).

@@

\textbf{\emph{Analysis Workflow:}} The total number of OTUs after
sub-sampling was analyzed for both the fecal and mock community samples.
Cycle dependent affects on Bray-Curtis indices were next assessed for
the fecal samples looking at both overall cycle differences and within
individual differences for the previous cycle (e.g.~20 versus 25, 25
versus 30, etc.). For these community based measures, the fecal samples
were analyzed at 4 different sub-sampling sequence depths (1000, 5000,
10000, and 15000) while the mock community samples were analysed at 3
levels (1000, 5000, 10000). Next, based on these observations we wanted
to analyze potential reasons for these differences. First, analysis of
general sequence error rate, number of sequences with an error, and base
substitution were assessed in the mock community for each DNA
polymerase. After assessing these errors, the total number of chimeras
was determined after sequence processing.

>>>>

\textbf{\emph{Human and Mock Samples:}} A single fecal sample was
obtained from 4 individuals who were part of the Enterics Research
Investigational Network (ERIN). The processing and storage of these
samples have been previously published (Seekatz et al., 2016). Other
than confirmation that none of these individuals had a \emph{Clostridium
difficle} infection, clinical data and other types of meta data were not
utilized or accessed for this study. All samples were extracted using
the MOBIO\textsuperscript{TM} PowerMag Microbiome RNA/DNA extraction kit
(now Qiagen, MD, USA). The ZymoBIOMICS\textsuperscript{TM} Microbial
Community DNA Standard (Zymo, CA, USA) was used in this study and is
made up of \emph{Pseudomonas aeruginosa}, \emph{Escherichia coli},
\emph{Salmonella enterica}, \emph{Lactobacillus fermentum},
\emph{Enterococcus faecalis}, \emph{Staphylococcus aureus},
\emph{Listeria monocytogenes}, and \emph{Bacillus subtilis} at equal
genomic DNA abundance
(\url{http://www.zymoresearch.com/microbiomics/microbial-standards/zymobiomics-microbial-community-standards}).

\textbf{\emph{PCR Protocol:}} The five different HiFi DNA polymerases
that were tested included AccuPrime\textsuperscript{TM} (ThermoFisher,
MA, USA), KAPA HIFI (Roche, IN, USA), Phusion (ThermoFisher, MA, USA),
Platinum (ThermoFisher, MA, USA), and Q5 (New England Biolabs, MA, USA).
The PCR cycle conditions for Platinum and Accuprime followed a
previously published protocol (Kozich et al., 2013)
(\url{https://github.com/SchlossLab/MiSeq_WetLab_SOP/blob/master/MiSeq_WetLab_SOP_v4.md}).
The HiFi DNA polymerase activation time was 2 minutes, unless a
different activation was specified. For Kappa and Q5, the protocol
previously published by Gohl and colleagues was used (Gohl et al.,
2016). For Phusion, the company defined conditions were used except for
extension time, where the Accuprime and Platinum settings were used.

@@

\textbf{\emph{Sequence Processing:}} The mothur software program was
utilized for all sequence processing steps (Schloss et al., 2009).
Generally, the protocol followed what has been previously published
(Kozich et al., 2013) (\url{https://www.mothur.org/wiki/MiSeq_SOP}). Two
major differences from the stated protocol were the use of VSEARCH
instead of UCHIME for chimera detection and the use of the OptiClust
algorithm instead of average neighbor for OTU generation at 97\%
similarity (Edgar et al., 2011; Rognes et al., 2016; Westcott \&
Schloss, 2017). Sequence error was determined using the `seq.error'
command on mock samples after chimera removal and classification to the
RDP to remove non-bacterial sequences (Schloss et al., 2009; Cole et
al., 2013; Rognes et al., 2016).

\textbf{\emph{Analysis Workflow:}} The total number of OTUs after
sub-sampling was analyzed for both the fecal and mock community samples.
Cycle dependent affects on Bray-Curtis indices were next assessed for
the fecal samples looking at both overall cycle differences and within
individual differences for the previous cycle (e.g.~20 versus 25, 25
versus 30, etc.). For these community based measures, the fecal samples
were analyzed at 4 different sub-sampling sequence depths (1000, 5000,
10000, and 15000) while the mock community samples were analysed at 3
levels (1000, 5000, 10000). Next, based on these observations we wanted
to analyze potential reasons for these differences. First, analysis of
general sequence error rate, number of sequences with an error, and base
substitution were assessed in the mock community for each HiFi DNA
polymerase. After assessing these errors, the total number of chimeras
was determined after sequence processing and correlations to the number
of OTUs also assessed.

@@
